Toxic Code Snippets on Stack Overflow

### 2018-05-24 cloverflow-tse/cloverflow.tex
Adopted the changes from the cover letter.

Although this activity of online code cloning is well-known, there is only a few
empirical studies on the topic~\cite{An2017,Abdalkareem2017,Baltes2017}, especially
on finding the origins of the clones on Q\&A websites.
In this study, we tackle this challenge of establishing existence of online code 
clones on Stack Overflow, investigate how they occur, and study the potential
effects to software reusing them.
%Using the same analogy, in this study, we are interested in 
Therefore, we mine Stack Overflow posts, detected online code clones, and analysed the clones to reveal
``toxic code snippets''.

@@

Nevertheless, we investigated a crucial parameter setting for clone detection: 
the minimum clone size threshold.
Choosing a large threshold value can reduce the number of trivial clones
(e.g.,~\texttt{equals}, \texttt{hashCode}, or getter and setter methods) and
false clones in the analysis or the manual investigation
phase~\cite{Sajnani2016}, while at the same time may create some false negatives,
i.e.,~increasing precision. On the other hand, setting a low threshold
results in a larger number of clone candidate pairs to look at, i.e.,~increasing recall, 
and a higher chance of getting false positives. 
%It is known that lowering the threshold value 
%is harmful to precision. 
Moreover, the large number of clone
pairs will hinders the full manual validation of the clones.
% and one needs to
%rely on sampling of clone pairs. 
Three threshold values, six, ten, and fifteen lines, were chosen for our
investigation. We started our investigation by using a threshold value of six
lines, a well-accepted minimum clone size in clone benchmark~\cite{Bellon2007}.
Simian reported 67,172 clone candidate pairs and SourcererCC reported 7,752
clone candidate pairs. We randomly sampled 382 pairs from the two sets for
manual check. This sample amount was a statistical significant sample with a
95\% confidence level and $\pm 5\%$ confidence interval. The first author
investigated the sampled clone pairs and classified them into three groups:
non-clones, trivial clones (\texttt{equals}, \texttt{hashCode}, or getter and
setter methods), and non-trivial clones. The manual check found 26 non-clone
pairs; 322 trivial clone pairs and 34 non-trivial clone pairs. Next, we increased the
threshold to ten lines, another well-established minimum clone size for
large-scale data sets~\cite{Sajnani2016}, and retrieved 721 clone pairs from
Simian and 1,678 clone pairs from SourcererCC. We randomly sampled and manually
checked the same amount of 382 pairs and found 27 non-clone pairs, 253 trivial
clone pairs and 102 non-trivial clone pairs. Then, we increased the threshold
further to fifteen lines and retrieved 196 clone pairs from Simian and 1,230
clone pairs from SourcererCC. The manual check of the 382 randomly sampled pairs
reveal zero non-clone pairs, 298 trivial clone pairs and 83 non-trivial clone
pairs.

@@

We did an analysis of the time it took for the clone to appear from Qualitas
projects to Stack Overflow. The clone ages were calculated by counting the
number of month difference from the dates of each Qualitas project to the date
the answers were posted on Stack Overflow as shown in~\ref{fig:boxplotcloneage}.
We found that, on average, it took the clones around 2 years since they appeared
in Qualitas projects to appear on Stack Overflow answers. Some of the clones
appeared on Stack Overflow almost at the same time as the original, while the
oldest clones took around 5 years.

@@

Then, we performed code clone detection between the 100 outdated code snippets and 130,703 GitHub projects.
We found 102 cloned candidates appearing in 68 GitHub projects and manually investigated all
of them. As shown in \Cref{tab:outdated_github}, out of the 102 cloned snippets, there were
13 cloned snippets that matched with themselves because some of the Qualitas
projects also appear on GitHub. For other projects besides the Qualitas
projects, there were 32 cloned snippets that were not exactly the same
(e.g.,~they contained additional code modifications made by the projects'
developers or they were copied from another source with a slightly different code).
There were 47 cloned snippets that were exactly the same as the
outdated code snippets, which of 12 were buggy.
Two cloned snippets gave attributions to Stack Overflow. The attributions
pointed to different posts than the ones we found, but containing the same code
in the answers. 32 cloned snippets were very likely to be a file-level
clone from its respective original project (e.g.,~JFreeChart, JUnit, Log4J,
Hadoop) based on their license header and the Javadoc comments. 
13 cloned snippets did not have any hints or evidence of copying.

@@

\subsubsection{Toxicity of outdated and license-incompatible clones}
We found only a small amount of toxic outdated code snippets in open source projects on
GitHub. Besides 12 buggy and outdated code snippets found in 12 projects, the
rest were non-harmful clones of the outdated code. Although other studies show
that Stack Overflow code snippets may become toxic by containing security
vulnerabilities~\cite{Acar2016,Fischer2017} or API misuse~\cite{Zhang2018}, we
found in this study that the damage caused by outdated code on Stack
Overflow is not high.

On the other hand, the missing of licensing statement of online code clones on
Stack Overflow can cause more damage than outdated code. As shown in our study and
also in the study by An et al. \cite{An2017}, some online clones on Stack Overflow are
originally licensed under more restrictive license than Stack Overflow's CC
BY-SA 3.0. If these missing-license online clones are reused in a software with
incompatible license, the software owner may face legal issues. The software
auditing service such as Black Duck Software or nexB, which can effectively
check for license compliance of code copied from open source projects, cannot
check for the original license of the cloned code snippets on Stack Overflow. Although
Stack Overflow answerers participated in our survey believe that most of the
code snippets on Stack Overflow are too small to claim for copyright and they
fall under fair-use, there is still a risk due to different legal system in each
country. For example, Germany legal system does not have a concept of fair use.
In addition, the number of minimum lines of code to be consider copying, i.e.,~de minimis,
is also differently interpreted from case to case or from country to country.

@@

In addition, we also performed a study of two open source software auditing
platforms/services:~BlackDuck
Software\footnote{https://www.blackducksoftware.com} and
nexB\footnote{https://www.nexb.com}. For BlackDuck Software, we found from their
report~\cite{CORSI2017} that while they check for code copied from open source
projects including GitHub and Stack Overflow and analyse their license
compatibility with their customer software, the BlackDuck auditing system will
treat the code snippets on Stack Overflow as having ``unknown'' license because
it does not know the original license of Stack Overflow code snippets. For nexB,
their product does not mention checking of reused source code from Stack
Overflow. So, our proposed service, which can offer more precise licensing
information of Stack Overflow code snippets, will be useful as an add-on license
check for code copying from Stack Overflow.

>>>>

Although this activity of online code cloning is well-known, there are only a few
empirical studies on the topic~\cite{An2017,Abdalkareem2017,Baltes2017}, especially
on finding the origins of the clones on Q\&A websites.
In this study, we tackle this challenge of establishing existence of online code 
clones on Stack Overflow, investigate how they occur, and study the potential
effects to software reusing them.
%Using the same analogy, in this study, we are interested in 
Therefore, we mine Stack Overflow posts, detected online code clones, and analysed the clones to reveal
``toxic code snippets''.

@@

Nevertheless, we investigated a crucial parameter setting for clone detection: 
the minimum clone size threshold.
Choosing a large threshold value can reduce the number of trivial clones
(e.g.,~\texttt{equals}, \texttt{hashCode}, or getter and setter methods) and
false clones in the analysis or the manual investigation
phase~\cite{Sajnani2016}, while at the same time may create some false negatives,
i.e.,~increasing precision. On the other hand, setting a low threshold
results in a larger number of clone candidate pairs to look at, i.e.,~increasing recall, 
and a higher chance of getting false positives. 
%It is known that lowering the threshold value 
%is harmful to precision. 
Moreover, the large number of clone
pairs will hinders the full manual validation of the clones.
% and one needs to
%rely on sampling of clone pairs. 
Three threshold values, six, ten, and fifteen lines, were chosen for our
investigation. We started our investigation by using a threshold value of six
lines, a well-accepted minimum clone size in clone benchmark~\cite{Bellon2007}.
Simian reported 67,172 clone candidate pairs and SourcererCC reported 7,752
clone candidate pairs. We randomly sampled 382 pairs from the two sets for
manual check. This sample number was a statistical significant sample with a
95\% confidence level and $\pm 5\%$ confidence interval. The first author
investigated the sampled clone pairs and classified them into three groups:
non-clones, trivial clones (\texttt{equals}, \texttt{hashCode}, or getter and
setter methods), and non-trivial clones. The manual check found 26 non-clone
pairs; 322 trivial clone pairs and 34 non-trivial clone pairs. Next, we increased the
threshold to ten lines, another well-established minimum clone size for
large-scale data sets~\cite{Sajnani2016}, and retrieved 721 clone pairs from
Simian and 1,678 clone pairs from SourcererCC. We randomly sampled and manually
checked the same amount of 382 pairs and found 27 non-clone pairs, 253 trivial
clone pairs and 102 non-trivial clone pairs. Then, we increased the threshold
further to fifteen lines and retrieved 196 clone pairs from Simian and 1,230
clone pairs from SourcererCC. The manual check of the 382 randomly sampled pairs
reveal zero non-clone pairs, 298 trivial clone pairs and 83 non-trivial clone
pairs.

@@

We did an analysis of the time it took for the clone to appear from Qualitas
projects to Stack Overflow. The clone ages were calculated by counting the
number of months between the date of each Qualitas project and the date the answer
was posted on Stack Overflow as shown in~\ref{fig:boxplotcloneage}.
We found that, on average, it took the clones around 2 years since they appeared
in Qualitas projects to appear on Stack Overflow answers. Some of the clones
appeared on Stack Overflow almost at the same time as the original, while the
oldest clones took around 5 years.

@@

Then, we performed code clone detection between the 100 outdated code snippets and 130,703 GitHub projects.
We found 102 cloned candidates, which were associated with 30 outdated code snippets, 
appearing in 68 GitHub projects and manually investigated all
of them. As shown in \Cref{tab:outdated_github}, out of the 102 cloned snippets, there were
13 cloned snippets that matched with themselves because some of the Qualitas
projects also appear on GitHub. For other projects besides the Qualitas
projects, there were 32 cloned snippets that were not exactly the same
(e.g.,~they contained additional code modifications made by the projects'
developers or they were copied from another source with a slightly different code).
There were 47 cloned snippets that were exactly the same as the
outdated code snippets, which of 12 were buggy.
Two cloned snippets gave attributions to Stack Overflow. The attributions
pointed to different posts than the ones we found, but containing the same code
in the answers. 32 cloned snippets were very likely to be a file-level
clone from its respective original project (e.g.,~JFreeChart, JUnit, Log4J,
Hadoop) based on their license header and the Javadoc comments. 
13 cloned snippets did not have any hints or evidence of copying.

@@

\subsubsection{Toxicity of outdated and license-incompatible clones}
We found only a small number of toxic outdated code snippets in open source projects on
GitHub. Besides 12 buggy and outdated code snippets found in 12 projects, the
rest were non-harmful clones of the outdated code. Although other studies show
that Stack Overflow code snippets may become toxic by containing security
vulnerabilities~\cite{Acar2016,Fischer2017} or API misuse~\cite{Zhang2018}, we
found in this study that the damage caused by outdated code on Stack
Overflow is not high.

On the other hand, the missing licensing statements of online code clones on
Stack Overflow can cause more damage than outdated code. As shown in our study and
also in the study by An et al. \cite{An2017}, some online clones on Stack Overflow are
originally licensed under more restrictive license than Stack Overflow's CC
BY-SA 3.0. If these missing-license online clones are reused in a software with
an incompatible license, the software owner may face legal issues. Software
auditing services such as Black Duck Software or nexB, which can effectively
check for license compliance of code copied from open source projects, do not
check for the original license of the cloned code snippets on Stack Overflow. Although
the Stack Overflow answerers who participated in our survey believe that most of the
code snippets on Stack Overflow are too small to claim for copyright and they
fall under fair-use, there is still a risk due to different legal systems in each
country. For example, Germany's legal system does not have a concept of fair use.
In addition, the number of minimum lines of code to be consider copying, i.e.,~de minimis,
is also differently interpreted from case to case or from country to country.

@@

In addition, we also performed a study of two open source software auditing
platforms/services:~BlackDuck
Software\footnote{https://www.blackducksoftware.com} and
nexB\footnote{https://www.nexb.com}. For BlackDuck Software, we found from their
report~\cite{CORSI2017} that while they check for code copied from open source
projects including GitHub and Stack Overflow and analyse their license
compatibility with their customer software, the BlackDuck auditing system will
treat the code snippets on Stack Overflow as having an ``unknown'' license because
it does not know the original license of Stack Overflow code snippets. For nexB,
their product does not mention checking of reused source code from Stack
Overflow. So, our proposed service, which can offer more precise licensing
information of Stack Overflow code snippets, will be useful as an add-on license
check for code copying from Stack Overflow.
